---
title: 'Reproducible Research: Peer Assessment 1'
author: "Ada Lee"
date: "Friday, January 16, 2015"
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---



## Loading and preprocessing the data
```{r readdata,echo=TRUE}
activity <- read.csv("activity.csv",colClasses = c("integer", "character", "character"))

format_interval <- function(c){
  res <- NA
  if(nchar(c) == 1) res <- paste("000", c, sep = "")
  if(nchar(c) == 2) res <- paste("00", c, sep = "")
  if(nchar(c) == 3) res <- paste("0", c, sep = "")
  if(nchar(c) == 4) res <- c
  res
}

activity$interval <- sapply(activity$interval, format_interval)
activity$date <- as.Date(activity$date, "%Y-%m-%d") # transform to Date
# look a bit of the data
head(activity)
tail(activity)
```

## What is mean total number of steps taken per day?
```{r totoal_numver_steps, echo=TRUE}
total_number_steps <- sapply(split(activity$steps, activity$date), sum, na.rm = TRUE)
mean_tns <- mean(total_number_steps)
median_tns <- median(total_number_steps)
dates <- as.Date(names(total_number_steps), "%Y-%m-%d")
hist(total_number_steps)
```

### The mean of total number of steps taken per day is `r mean_tns`.  
### The median of total number of steps taken per day is `r median_tns`.

## What is the average daily activity pattern?
```{r avg_daily_pattern, echo=TRUE}
avg_daily <- sapply(split(activity$steps, activity$interval), mean, na.rm = TRUE)
hms <- strptime(names(avg_daily), "%H%M")
plot(hms, avg_daily, type = "l", main = "Average Daily Activity Pattern",  
     xlab = "interval", ylab = "steps")
max_avg_daily <- max(avg_daily)
avg_daily <- data.frame(interval = names(avg_daily), avg = avg_daily, stringsAsFactors = FALSE)
res <- subset(avg_daily, avg == max_avg_daily)
```
### `r res$interval` of 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps (`r res$avg`).

## Imputing missing values
```{r missing_values, echo=TRUE}
total_missing <- sum(is.na(activity$steps))
filled_activity <- activity

n <- length(filled_activity$steps)

# filling in all of the missing values of steps, use the mean for that 5-minute interval
for (i in 1:n){
  item <- filled_activity[i, ]
  if (is.na(item$steps)){
    imputting <- subset(avg_daily, interval == item$interval)$avg
    filled_activity[i, ]$steps <- imputting
  }
}

# What is mean total number of steps taken per day after imputting missing values
total_number_steps <- sapply(split(filled_activity$steps, filled_activity$date), sum, na.rm = TRUE)
mean_tns <- mean(total_number_steps)
median_tns <- median(total_number_steps)
dates <- as.Date(names(total_number_steps), "%Y-%m-%d")
# histgram of total_number_steps after filling missing value
hist(total_number_steps, main = "histgram of total_number_steps after filling missing value")  
```

### After filling missing values, the mean of total number of steps taken per day is `r mean_tns`. And the median of total number of steps taken per day is `r median_tns`. These values differ from the estimates from the first part of the assignment. 

## Are there differences in activity patterns between weekdays and weekends?
```{r patterns}
weekday_or_weekend <- function(d){
  weekend <- c("Saturday", "Sunday")
  weekdays_d <- weekdays(d)
  if (weekdays_d %in% weekend){res <- "weekend"}
  else {res <- "weekday"}
  res
}

# create a new factor variable in the dataset with two levels -"weekday" 
# and "weekend" indicating whether a given date is a weekday or weekend day
filled_activity$weekday_or_weekend <- as.factor(sapply(filled_activity$date, weekday_or_weekend))
# summary filled_activity
summary(filled_activity)

splited_activity <- split(filled_activity, filled_activity$weekday_or_weekend)
weekday_activity <- splited_activity[["weekday"]]
weekend_activity <- splited_activity[["weekend"]]

## weekday 
weekday_avg_daily <- sapply(split(weekday_activity$steps, weekday_activity$interval),mean)
weekday_hms <- strptime(names(weekday_avg_daily), "%H%M")

## weekend
weekend_avg_daily <- sapply(split(weekend_activity$steps, weekend_activity$interval),  mean)
weekend_hms <- strptime(names(weekend_avg_daily), "%H%M")

## 
par(mfrow = c(2, 1))
ylim <- c(0, ceiling(max(weekday_avg_daily, weekend_avg_daily)))

## plot weekday
plot(weekday_hms, weekday_avg_daily, type = "l", main = "Average Daily Activity Pattern on Weekday",  
     ylim=ylim,   xlab = "interval", ylab = "steps")
# plot on weekend
plot(weekend_hms, weekend_avg_daily, type = "l", main = "Average Daily Activity Pattern on Weekend",  
     ylim=ylim,   xlab = "interval", ylab = "steps")
```

### Weekdays have higher max steps.
