---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document:
    keep_md: yes
  pdf_document: default
  word_document: default
---


## Loading and preprocessing the data
```{r readdata,echo=TRUE}
activity <- read.csv("activity.csv",colClasses = c("integer", "character", "character"))

format_interval <- function(c){
  res <- c
  if(nchar(c) == 1) res <- paste("000", c, sep = "")
  if(nchar(c) == 2) res <- paste("00", c, sep = "")
  if(nchar(c) == 3) res <- paste("0", c, sep = "")
  if(nchar(c) == 0) res <- NA
  
  res
}

activity$interval <- sapply(activity$interval, format_interval)
activity$date <- as.Date(activity$date, "%Y-%m-%d") # transform to Date
# look a bit of the data
head(activity)
tail(activity)
```

## What is mean total number of steps taken per day?
```{r totoal_numver_steps, echo=TRUE}
total_number_steps <- sapply(split(activity$steps, activity$date), sum, na.rm = TRUE)
mean_tns <- mean(total_number_steps)
median_tns <- median(total_number_steps)
dates <- as.Date(names(total_number_steps), "%Y-%m-%d")
#plot(dates, total_number_steps, type = "h", main = "Total Number of Steps Taken Each Day (2012)")
hist(total_number_steps)
```

The mean of total number of steps taken per day is `r mean_tns`.  
The median of total number of steps taken per day is `r median_tns`.

## What is the average daily activity pattern?
```{r avg_daily_pattern, echo=TRUE}
avg_daily <- sapply(split(activity$steps, activity$interval), mean, na.rm = TRUE)
hms <- strptime(names(avg_daily), "%H%M")
plot(hms, avg_daily, type = "l", main = "Average Daily Activity Pattern")
max_avg_daily <- max(avg_daily)
avg_daily <- data.frame(interval = names(avg_daily), avg = avg_daily, stringsAsFactors = FALSE)
res <- subset(avg_daily, avg == max_avg_daily)
```
`r res$interval` 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps (`r res$avg`).

## Imputing missing values
```{r missing_values, echo=TRUE}
total_missing <- sum(is.na(activity$steps))
new_activity <- activity

n <- length(new_activity$steps)

# filling in all of the missing values of steps
# use the mean for that 5-minute interval
for (i in 1:n){
  item <- new_activity[i, ]
  if (is.na(item$steps)){
    imputting <- subset(avg_daily, interval == item$interval)$avg
    new_activity[i, ]$steps <- imputting
  }
}

# What is mean total number of steps taken per day
# after imputting missing values
total_number_steps <- sapply(split(new_activity$steps, new_activity$date), sum, na.rm = TRUE)
mean_tns <- mean(total_number_steps)
median_tns <- median(total_number_steps)
dates <- as.Date(names(total_number_steps), "%Y-%m-%d")
#plot(dates, total_number_steps, type = "h", main = "Total Number of Steps Taken Each Day (2012)")
hist(total_number_steps, main = "histgram of total_number_steps after filling missing value")


```
<br></br>
After filling missing values, the mean of total number of steps taken per day is `r mean_tns`. And the median of total number of steps taken per day is `r median_tns`. These values differ from the estimates from the first part of the assignment. 

## Are there differences in activity patterns between weekdays and weekends?
